---
title: "Systematicity across grammatical category and other lexical features"
output: html_document
---

# Introduction

This report investigates whether the **sub-morphemic form-meaning systematicity** of different words correlates with various lexical features, including grammatical category, age of acquisition, and word frequency.

Previous work ([Monaghan et al, 2014](http://rstb.royalsocietypublishing.org/content/369/1651/20130299.short)) has found that age of acquisition (AoA) predicts systematicity, above and beyond other variables like orthographic/phonological length. Specifically, earlier-acquired words were more systematic.

# Dataset

```{r}
setwd("/Users/strott/nlp/systematicity/")
```


## Systematicity of individual words

```{r}
setwd("/Users/strott/nlp/systematicity/")
word_systematicity_scores = read_csv("data/processed/all_words_systematicity.csv")
```

We calculated the systematicity of each word following the same procedure as [Monaghan et al, 2014](http://rstb.royalsocietypublishing.org/content/369/1651/20130299.short). The Jupyter notebook containing this analysis can be found [here](https://github.com/seantrott/systematicity/blob/master/systematicity_by_word_class.ipynb). 
We obtained monosyllabic, monomorphemic words from CELEX, then filtered on the words that appeared in the Google News word embeddings, resulting in `r nrow(word_systematicity_scores)` words overall.

We then calculated a **global correlation** score between the *form* and *meaning* distances of each wordpair. To obtain by-word systematicity scores, we compared this global correlation to the correlation coefficient when omitting a particular word: the difference in scores is deemed the *contribution* a given word makes to a lexicon's global systematicity.

The distribution of systematicity scores is plotted below:

```{r}
word_systematicity_scores %>%
  ggplot(aes(x = impact)) +
  geom_histogram() +
  labs(x = "Contribution to systematicity",
       title = "Distribution of systematicity scores") +
  theme_minimal()
```

## Age of Acquisition data

```{r}
setwd("/Users/strott/nlp/systematicity/")
aoa_kuperman = read_csv("data/raw/AoA_Kuperman.csv")
aoa_kuperman$word = aoa_kuperman$Word
```

For AoA data, we used the same dataset that [Monaghan et al, 2014](http://rstb.royalsocietypublishing.org/content/369/1651/20130299.short) used, collected by [Kuperman et al (2012)](http://crr.ugent.be/archives/806).

There are scores for `r nrow(aoa_kuperman)` words overall.

```{r}
aoa_kuperman %>%
  ggplot(aes(x = Rating.Mean)) +
  geom_histogram() +
  labs(x = "Mean AoA Rating",
       title = "Distribution of AoA ratings") +
  theme_minimal()
```

## Word class data

```{r}
setwd("/Users/strott/nlp/systematicity/")
word_classes = read.csv("data/raw/celex_classes.csv", sep="\\")
word_classes$word = word_classes$Word
```

Finally, we used [WebCelex](http://celex.mpi.nl/) to grab information about the word class and frequency of all monomorphemic, monosyllabic words. There is information for `r nrow(word_classes)` overall.

```{r}
table(word_classes$Class)
```

As [Monaghan et al 2014](http://rstb.royalsocietypublishing.org/content/369/1651/20130299.short) do, we remove words with more than one word class:

```{r}
word_classes$duplicate = duplicated(word_classes$Word)

duplicate_words = filter(word_classes, duplicate==TRUE)$Word
word_classes_filtered = filter(word_classes, (Word %in% duplicate_words) == FALSE)

```

## Combining all together

Finally, we merged each of these datasets together:

```{r}
merged = merge(word_systematicity_scores, aoa_kuperman, on="word")
merged_all = merge(merged, distinct(word_classes_filtered), on="word")
```

This results in only `r nrow(merged_all)` observations.

# Results

Our chief question is whether the **systematicity** of a word is related to its:

* Grammatical Class
* Age of Acquisition
* Frequency
* Length

To answer this, we regressed *systematicity* in a standard OLS regression against each of those variables:

```{r}
model_omnibus = lm(data = merged_all,
                   impact ~ Rating.Mean + Class + CobLog + word_length)
summary(model_omnibus)
```

Unlike Monaghan et al (2014), we don't see a significant relationship between AoA and systematicity. This is also evident if we visualize that relationship:

```{r}
merged_all %>%
  ggplot(aes(x = round(Rating.Mean),
             y = impact)) +
  geom_line(stat = "summary", fun.y = "mean") +
  theme_minimal()
```


# Next Steps

In order, it seems like we ought to:

* Do further validation of the dataset, to account for the divergence in results from Monaghan et al, 2014.
* Look at other lexical features.
* Look across languages.
